---
title: "Gender Equity"
author: "Ama Owusu-Darko"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load R packages

```{r load_r_packages}
library(here)
library(dplyr)
library(ggplot2)
library(readxl)
library(tidyr)
library(stringr)
```

Setup up file paths


```{r file_setup}
here::i_am("analysis/Ama_K_Owusu_Darko_Assignment.Rmd")
```

Load Pre-work data set and save it as the dataframe "school_leaders_df

```{r load_dat}
school_leaders_df <- readxl::read_excel(here::here("raw_data/Pre-work Dataset.xlsx"))


```

Data Preparation
```{r subset}

#Creating a subset of the data
professions_df <- school_leaders_df[,c(1,8,42:50)] |> 
  tidyr::pivot_longer(
    cols = starts_with("What kinds of careers do you think are suitable for students?"),
    names_to = "Professions",
    values_to = "Suitable_Gender"
  )


# Cleaning up names of profession
professions_df$Professions <- stringr::str_replace(professions_df$Professions,"What kinds of careers do you think are suitable for students?", "" )

professions_df$Professions <- stringr::str_replace_all(professions_df$Professions,"[?\\n ]", "" )


#Separating data set for the school leader is a man 

professions_df_male_school_leader <- professions_df |>  
  dplyr::mutate(dplyr::across(c(1:3), as.factor)) |> 
  dplyr::filter(`What is your gender?` == "Man") |> 
  dplyr::count(`What is your gender?`,Professions, Suitable_Gender, name = "Count") |> 
  dplyr::arrange(Professions) |> 
  dplyr::mutate(Within_Group_percentage = round((Count/ 302)*100, 1))


#Separating data set for the school leader is a woman
professions_df_female_school_leader <- professions_df |>  
  dplyr::mutate(dplyr::across(c(1:3), as.factor)) |> 
  dplyr::filter(`What is your gender?` == "Woman")|>  
  dplyr::count(`What is your gender?`, Professions, Suitable_Gender, name = "Count") |> 
  dplyr::arrange(Professions) |> 
  dplyr::mutate(Within_Group_percentage = round((Count/ 183)*100, 1))


professions_df_female_school_leader
professions_df_male_school_leader
```

### Creating a Gender versus Professions Percent stacked bar chart 
To display the results of which genders the school leaders perceive to be suitable for the different professions. 

```{r bargraph}

ggplot2::ggplot(professions_df_male_school_leader, aes(fill = Suitable_Gender,  x = Professions, y=Within_Group_percentage ))+
    geom_bar(position="stack", stat="identity")+
scale_fill_manual(values = c("#e8e9eb", "#66aa00", "#808", "#e8e"))+
  theme_minimal()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
 labs(title="Opinions of Male School leaders",  subtitle = "Figure 1. On suitable professions for Boys versus Girls") + annotate("text", x = c(1,3, 6), y = c(30), 
label = c( paste(professions_df_male_school_leader[3,5],"%"), 
          paste(professions_df_male_school_leader[11,5],"%"),
          paste(professions_df_male_school_leader[23,5],"%")), color="white", 
size=3 , angle=0, fontface="bold")+
  annotate("text", x = c(4,8), y = c(80), 
label = c( paste(professions_df_male_school_leader[13,5],"%"),
          paste(professions_df_male_school_leader[25,5],"%")), color="black", 
size=3 , angle=0, fontface="bold")

ggplot2::ggplot(professions_df_female_school_leader, aes(fill = Suitable_Gender,  x = Professions, y= Within_Group_percentage ))+
    geom_bar(position="stack", stat="identity")+
scale_fill_manual(values = c("#e8e9eb", "#66aa00", "#808", "#e8e"))+
  theme_minimal()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title="Opinions of Female School leaders",  subtitle = " Figure 2. On suitable professions for Boys versus Girls")+ annotate("text", x = c(1,3, 6), y = c(30), 
label = c( paste(professions_df_female_school_leader[3,5],"%"), 
          paste(professions_df_female_school_leader[11,5],"%"),
          paste(professions_df_female_school_leader[23,5],"%")), color="white", 
size=3 , angle=0, fontface="bold")+
  annotate("text", x = c(4,8), y = c(80), 
label = c( paste(professions_df_female_school_leader[13,5],"%"),
          paste(professions_df_female_school_leader[25,5],"%")), color="black", 
size=3 , angle=0, fontface="bold")
```

The survey results showed that more female than male school leaders have segregated ideas about which careers are suited for each gender. For instance, as shown in Figures 1 and 2 above, the difference is skewed towards Female School leaders thinking girls are more suited to being Airhostesses (61.7%) versus 47.7% of Male School leaders. 

The following results also support the point that female school leaders may be more inclined to think certain professions are for boys rather than being suited to both genders.Only 60.7% of female school leaders perceive that both girls and boys are suited to be doctors compared to 71.5% of male school leaders.


## What is the effect of school leaders being trained on gender within the last year?.

The profession Airhostess would be referred to as Flight attendant from now on to be more inclusive right from the name.

Focusing on the professions Flight Attendant and Pilot

```{r air_professions}
#Separating data set for the school leader is a man whole receive gender training and the assessed professions are Air Hostess and Pilot

air_df_male_school_leader <- professions_df |>  
  dplyr::filter(Professions %in% c("[Airhostess]", "[Pilot]")) |> 
  dplyr::filter(`Have you received any training related to gender within the last 1 year?` == "Yes") |> 
  dplyr::mutate(dplyr::across(c(1:3), as.factor)) |> 
  dplyr::filter(`What is your gender?` == "Man") |> 
  dplyr::count(`What is your gender?`,`Have you received any training related to gender within the last 1 year?`,Professions, Suitable_Gender, name = "Count") |> 
  dplyr::arrange(Professions) |> 
  dplyr::mutate(Within_Group_percentage = round((Count/ 302)*100, 1)) # There are 302 males school leaders


#Separating data set for the school leader is a woman and the assessed professions are Air Hostess and Pilot
air_df_female_school_leader <- professions_df |>  
  dplyr::filter(Professions %in% c("[Airhostess]", "[Pilot]")) |>
  dplyr::filter(`Have you received any training related to gender within the last 1 year?` == "Yes") |>
  dplyr::mutate(dplyr::across(c(1:3), as.factor)) |> 
  dplyr::filter(`What is your gender?` == "Woman")|>  
  dplyr::count(`What is your gender?`,`Have you received any training related to gender within the last 1 year?`,Professions, Suitable_Gender, name = "Count") |> 
  dplyr::arrange(Professions) |> 
  dplyr::mutate(Within_Group_percentage = round((Count/ 183)*100, 1)) # There are 183 female school leaders


air_professions_df <- rbind(air_df_female_school_leader, air_df_male_school_leader)


# Bar graph
ggplot2::ggplot(air_professions_df, aes(fill =  `What is your gender?`,  x = Suitable_Gender, y= Count ))+
    geom_bar(position="dodge", stat="identity")+
scale_fill_manual(values = c( "#66aa00", "#808"))+
  theme_minimal()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
 labs(title="Opinions of Male School leaders",  subtitle = "Figure 1. On suitable professions for Boys versus Girls")+
  facet_wrap(~Professions)
```


are school leaders who have received gender training in the last year more likely to say 